---
title: "tbl_df with S4 Objects"
date: "`r format(Sys.time(), '%d %B, %Y')`"
---

```{r}
library("knitr")
opts_chunk$set(error = TRUE)
```

# Background 

This is my first attempt at doing object oriented R programming. Specifically, I've declared a S4 object, TestObject, that has two slots where one of the slots (test.df) has the permitted class of a data.frame. 

```{r, message = FALSE}
library("dplyr")
library("methods")

# Declare S4 Object
setClass("TestObject",
  slots = list(test.mat = "matrix",
               test.df = "data.frame"))
```

To my surprise, when I try to assign an object of tbl_class "class":

```{r}
library("nycflights13")
class(flights)
```

It throws an error:

```{r}
test.object <- new("TestObject")
slot(test.object, "test.mat") <- matrix(rnorm(10), 5, 2)
slot(test.object, "test.df") <- flights
```

I was under the impression that the tbl_df just wraps around a data.frame. I thought maybe I could declare the tbl_df as the permitted class for test.df:

```{r}
setClass("TestObject",
  slots = list(test.mat = "matrix",
               test.df = "tbl_df"))
```

It seems that I would have to convert the tbl_df into a data.frame, in order for this work...

```{r}
setClass("TestObject",
  slots = list(test.mat = "matrix",
               test.df = "data.frame"))

slot(test.object, "test.df") <- as.data.frame(flights)
head(slot(test.object, "test.df"))
```
